
---
title: "Dummy Data"
author: "Marit Kuijt"
date: "2023-02-14"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/marit/OneDrive/Documenten/Masterstage Naturalis/brassicaceae-trait-climate-analysis")
```

```{r, include=FALSE}
library(geodata)
options(stringsAsFactors = FALSE)
library(raster)
library(tidyverse, quietly = TRUE)
library(sp)
library(adehabitatMA)
library(optparse)
library(yaml)
library(dplyr)
library(ade4)
library(tibble)
library(logger)
library(CoordinateCleaner)
library(fuzzySim)
library(factoextra)
library(ggforce)
library(ggalt)
library(ggtree)
library(usdm)
library(subniche)
library(knitr)
library(phylolm)
library(ggnewscale)
library(ggplot2)
library(ggpubr)
library(ape)
library(phytools)
library(maptools)
library(Hmisc)
library(RRphylo)
```
#Read Data
##GIS data
```{r add_layers_ENVIREM}
# Read ENVIREM data and stack
remove(WorldENVIREM)
WorldENVIREM <- raster()

# The location of the TIFF file with the stacked layers described directly above
files.names <- list.files("./Data/ENVIREM")
# Turn the file names into layer names: strip the prefix (which might include
# the resolution) and strip the file extension
gis.layers.names <- files.names
gis.layers.names <- gsub('./Data/wc2.1_10m_','',gis.layers.names)
gis.layers.names <- gsub('.tif','',gis.layers.names)

gis.layers <- c()
# Iterate over files
for (j in 1:length(files.names)) {
  
  # Stack with previously read layers
  WorldENVIREM <- stack(
    WorldENVIREM,
    
    # Read as raster
    raster(
      
      # Construct file name
      paste("./Data/ENVIREM/", files.names[j], sep = "")
    )
  )
}
```

##UV Data
```{r}
UV_Data <- read.asciigrid("./Data/56459_UVB1_Annual_Mean_UV-B.asc")
plot(UV_Data)
UV_Data.raster <- raster(UV_Data)
sp::proj4string(UV_Data) <- crs("+proj=longlat +datum=WGS84")
```

###Extract coordinates from cells of ENVIREM and values from these cells of UV data
```{r}
coordinates <- raster::xyFromCell(WorldENVIREM, 1:1944000)
UV_cells <- raster::cellFromXY(UV_Data.raster, coordinates)
UV_values <- raster::extract(UV_Data.raster, UV_cells)
UV_Bound <- cbind(coordinates, UV_values)
UV_Bound <- data.frame(UV_Bound)
sp::coordinates(UV_Bound) <- ~ x + y
sp::proj4string(UV_Bound) <- crs("+proj=longlat +ellps=WGS84")

UV_Bound.raster <- rasterFromXYZ(UV_Bound)
sp::proj4string(UV_Bound.raster) <- crs("+proj=longlat +ellps=WGS84")
```

```{r}
WorldENVIREM <- stack(WorldENVIREM, UV_Bound.raster)

rm(gis.layers.names, files.names, j)
```


##BioClim Data
```{r add_layers_World_clim}
## Read Bioclim data and stack
remove(WorldBioClim)
WorldBioClim <- raster()

# The location of the TIFF file with the stacked layers described directly above
files.names <- list.files("./Data/wc2.1_10m")
# Turn the file names into layer names: strip the prefix (which might include
# the resolution) and strip the file extension
gis.layers.names <- files.names
gis.layers.names <- gsub('./Data/wc2.1_10m_','',gis.layers.names)
gis.layers.names <- gsub('.tif','',gis.layers.names)

gis.layers <- c()
# Iterate over files
for (i in 1:length(files.names)) {
  
  # Stack with previously read layers
  BioClimLayer <- raster(paste("./Data/wc2.1_10m/", files.names[i], sep = ""))
  BioClimLayer <- crop(BioClimLayer, WorldENVIREM)
  WorldBioClim <- stack(
    WorldBioClim, BioClimLayer
  )
}

rm(gis.layers.names, files.names, i)
```

##Stack all BioClimatic Layers together
```{r}
All_Layers <- stack(WorldBioClim, WorldENVIREM)
All_Layers.spdf <- as(All_Layers, "SpatialPixelsDataFrame")
sp::proj4string(All_Layers.spdf) <- crs("+proj=longlat +ellps=WGS84")
```

##Generate random data
```{r}
Mock_Species <- paste("species", 1:40, sep = "")
Mock_Species

for(i in 1:length(Mock_Species)){
  if (i < 20){
    Seqlat <- seq(from=15, to=25, by=.01)
    Seqlong <- seq(from=-6, to=25, by=.01)
    Latitude <- sample(Seqlat, size=100, replace=TRUE)
    Longitude <- sample(Seqlong, size=100, replace=TRUE)
    Growth_State <- replicate(100, 1)
    VirtualCoordinates <- data.frame(x= Longitude, y = Latitude, species = Mock_Species[i], Growth_State)
    write.csv(VirtualCoordinates, file = paste("Data/MockData/", Mock_Species[i], ".csv", sep = ""), quote = F)
  }
  else {
    Seqlat <- seq(from=-11, to=2, by=.01)
    Seqlong <- seq(from=-74, to=-53, by=.01)
    Latitude <- sample(Seqlat, size=100, replace=TRUE)
    Longitude <- sample(Seqlong, size=100, replace=TRUE)
    Growth_State <- replicate(100, 0)
    VirtualCoordinates <- data.frame(x= Longitude, y = Latitude, species = Mock_Species[i], Growth_State)
    write.csv(VirtualCoordinates, file = paste("Data/MockData/", Mock_Species[i], ".csv", sep = ""), quote = F)
  }  
}
```


##Occurrence data
```{r}
occurrences.spdf <- new(
    "SpatialPointsDataFrame", 
    coords = structure(
        numeric(0), 
        .Dim = c(0L, 2L),
        .Dimnames = list( NULL, c("x", "y") )
    ),  
    bbox = structure(
        c(1,1,1,1), 
        .Dim = c(2L, 2L),                         
        .Dimnames = list( c("x","y"), c("min","max") )
    )
) 
sp::proj4string(occurrences.spdf) <- crs("+proj=longlat +ellps=WGS84")
```

```{r}
for (j in 1:length(Mock_Species)) {
    
    # Prevent out of bounds errors
    if (j > length(Mock_Species)) {
        break
    }
  
	# Read occurrences
    csv.file <- paste("Data/MockData/", Mock_Species[j], ".csv", sep = "")
    csv.data <- read.csv(csv.file)
    #tryCatch(csv.data <- read.csv(csv.file), error=function(e) force(do.next))
    points.df <- dplyr::select(
        csv.data,
        x,
        y,
        species
    )
    #Coordinate Cleaner
    points.df <- na.omit(points.df)
    #points.df <- points.df[rowSums(points.df[,1:2]) > 0,]

    
    sp::coordinates(points.df) <- ~ x + y
    sp::proj4string(points.df) <- crs("+proj=longlat +ellps=WGS84")
    
    # Populate SpatialPointsDataFrame for focal taxon
    focal.spdf <- SpatialPointsDataFrame(
    	points.df, 
    	data.frame(species = rep(Mock_Species[j], NROW(points.df)))
    )
    sp::proj4string(focal.spdf) <- crs("+proj=longlat +ellps=WGS84")
    
    # Append to cumulative data frame
    occurrences.spdf <- rbind(occurrences.spdf, focal.spdf)
    
}
```


###Plot occurrences
```{r}
occurrences.df <- data.frame(occurrences.spdf@data)
occurrences.df <- cbind(occurrences.df, occurrences.spdf@coords)
occurrences.df$Growth_State <- 0
occurrences.df$Growth_State[0:1900] <- 1
lon <- dplyr::select(occurrences.df, x)
lat <- dplyr::select(occurrences.df, y)

data("wrld_simpl")
shape <- wrld_simpl
shape_df <- fortify(shape, region = 'ISO3')

ggplot() +
  geom_polygon(data = shape_df, aes(x = long, y = lat, group = group), fill = "white", color = "black")+
  geom_point(aes(x, y, fill = Growth_State), data = occurrences.df, shape = 21, size = 2, alpha = 0.5, col = "black") +
    coord_equal(xlim = c((min(lon) - 1), (max(lon) + 1)), ylim = c(min(lat) - 1, max(lat) + 1)) +
  xlab("Longitude") + ylab("Latitude") +
  theme(panel.border = element_rect(fill=NA), legend.position = 'right')
```


```{r}
names <- unique(occurrences.spdf$species)
```

##GIS data per cell (na.omit to exclude areas without data in oceans etc.)
```{r}
traitvalues.df <- slot(All_Layers.spdf, "data")
traitvalues.df <- tibble::rowid_to_column(traitvalues.df, "ID")
traitvalues.df <- na.omit(traitvalues.df)
traitvalues.df$UV <- traitvalues.df$layer
traitvalues.df <- subset(traitvalues.df, select = -layer)
```

##Presence/absence matrix per cell
```{r}
npoints <- adehabitatMA::count.points(occurrences.spdf, All_Layers.spdf)
noccurrences.df <- slot(npoints, "data")
noccurrences.df <- tibble::rowid_to_column(noccurrences.df, "ID")
noccurrences.df <- subset(noccurrences.df, noccurrences.df$ID %in% traitvalues.df$ID)
noccurrences.df <- subset(noccurrences.df, select = -ID)
noccurrences.df[noccurrences.df > 0] <- 1
noccurrences.df <- noccurrences.df[rowSums(noccurrences.df[]) > 0,]
```

##Select only cells with presence for noccurrences.df and traitvalues.df
```{r}
noccurrences.df$ID <- rownames(noccurrences.df)
traitvalues.df <- subset(traitvalues.df, traitvalues.df$ID %in% noccurrences.df$ID)
noccurrences.df <- subset(noccurrences.df, select = -ID)
```


#OMI analysis
##First a PCA
```{r}
traitvalues.dudi <- ade4::dudi.pca(
    traitvalues.df[2:length(traitvalues.df)], 
    scannf = F
)
scatter(traitvalues.dudi)
```

##Then the OMI analysis
```{r}
niche.dudi <- ade4::niche(traitvalues.dudi, noccurrences.df, scannf = F)
par(mfrow = c(1,1))
plot(niche.dudi)
screeplot(niche.dudi)
kable(niche.param(niche.dudi))
rtest(niche.dudi)
```

```{r}
rbind(
  Eigenvalue = niche.dudi$eig,
  Proportion = niche.dudi$eig/sum(niche.dudi$eig),
  Cumulative = cumsum(niche.dudi$eig)/sum(niche.dudi$eig))
```

##Save output
```{r}
write.csv(niche.dudi$tab, file = "./Results/niche_traits_merged_Lepidium", append = T, quote = F )
```


#Read Tree and filter to species in OMI
```{r}
set.seed(184)
Phylogeny_clustered <- ape::rtree(40)
Phylogeny_clustered$tip.label <- Mock_Species
plot(Phylogeny_clustered)
```

```{r}
set.seed(184)
Phylogeny_random <- ape::rtree(40, tip.label = Mock_Species)
plot(Phylogeny_random)
```


#Combine trait data of species and enviroment
```{r}
MeanBioClim <- niche.dudi$tab
Growth_Form <- data.frame(Mock_Species)
rownames(Growth_Form) <- Growth_Form$Mock_Species
Growth_Form$Growth_habit <- 1
Growth_Form$Growth_habit[20:40] <- 0
```

```{r}
Growth_Form$Growth_habit <- base::as.numeric(Growth_Form$Growth_habit)
```

##Add the ENVIREM OMI-transformed climatic data for each species
```{r}
Growth_Form[,3:38] <- NA
Trait_names <- colnames(MeanBioClim)
colnames(Growth_Form)[3:38] <- Trait_names
for (i in 1:length(Trait_names)) {
  for(j in 1:length(Mock_Species)){
    Growth_Form[Growth_Form$Mock_Species == Mock_Species[j], colnames(Growth_Form) == Trait_names[i]] <- MeanBioClim[rownames(MeanBioClim) == Mock_Species[j], colnames(MeanBioClim) == Trait_names[i]]
  }
}
Growth_Form <- Growth_Form[,c(1:3, 14:21, 4:13, 22:38)]
```

```{r}
Growth_Form <- na.omit(Growth_Form)
```

```{r}
Growth_Form_Name <- subset(Growth_Form, select = Mock_Species)
Growth_Form_Trait <- subset(Growth_Form, select = Growth_habit)
Growth_Form_Trait$Growth_habit <- as.factor(Growth_Form_Trait$Growth_habit)
Growth_Form_UV <- subset(Growth_Form, select = UV)
```

```{r}
Growth_Form$Growth_habit <- factor(Growth_Form$Growth_habit, levels = 0:1, labels = c(0,1))
```
#Plot tree
```{r}
p <- ggtree(Phylogeny_clustered) %<+% Growth_Form +
  geom_tiplab(size = 3)
  #geom_tiplab(aes(label = Growth_Form_Name), offset = 2, size = 3)
p2 <- gheatmap(p, Growth_Form_Trait, offset = 5, width = 0.1)
p3 <- p2 + new_scale_fill()
gheatmap(p3, Growth_Form_UV, offset = 8, width = 0.1) + scale_fill_continuous(na.value = "white")

p <- ggtree(Phylogeny_random) %<+% Growth_Form +
  geom_tiplab(size = 3)
  #geom_tiplab(aes(label = Growth_Form_Name), offset = 2, size = 3)
p2 <- gheatmap(p, Growth_Form_Trait, offset = 5, width = 0.1)
p3 <- p2 + new_scale_fill()
gheatmap(p3, Growth_Form_UV, offset = 8, width = 0.1) + scale_fill_continuous(na.value = "white")

```

#Boxplots
##Make boxplots per variable ENVIREM

```{r}
Growth_Form_Long <- Growth_Form
Growth_Form_Long$ID <- rownames(Growth_Form_Long)
Growth_Form_Long <- Growth_Form_Long[,c(1:2, 39, 3:38)]
Growth_Form_Long <- gather(Growth_Form_Long, BCVariable, Value, 4:39, factor_key = T)

boxplot <- ggplot(Growth_Form_Long, aes(Growth_habit, Value, group = Growth_habit, fill = Growth_habit)) +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y.npc = "center") +
facet_wrap(~ BCVariable)
boxplot
```

#Make a dendrogram of all variables to see the relationship between them

```{r, warning=FALSE}
MeanBioClim.matrix <- as.matrix(MeanBioClim)
MeanBioClim.matrix <- scale(MeanBioClim.matrix)
v <- varclus(MeanBioClim.matrix)
h <- v$hclust
hd <- as.dendrogram(h)
dend <- fviz_dend(hd, k = 5, cex=0.8, palette = "jco", color_labels_by_k = T, 
      rect = T, rect_lty = 0, rect_border = "jco", rect_fill = T)
dend
vars <- dend$layers[[2]]$data
clusters <- dend$layers[[4]]$data

colours <- unique(vars$col)
randomvars <- data.frame()
variables <- c()

for (i in 1:length(colours)) {
showvar <- vars[sample(which(vars$col == colours[i]), 2), ]
randomvars <- rbind(randomvars, showvar)
}
randomvars$label
```



#Make a model to use in phyloglm based on the boxplots and vif
```{r}
unique(Growth_Form$Name)

formula_1 <- Growth_habit ~  wc2.1_10m_bio_1 + wc2.1_10m_bio_8 + current_10arcmin_PETDriestQuarter + current_10arcmin_PETWarmestQuarter + wc2.1_10m_bio_5 + current_10arcmin_PETWarmestQuarter + wc2.1_10m_bio_2 + wc2.1_10m_bio_15 + wc2.1_10m_bio_13 + wc2.1_10m_bio_11       
```

#Make brench lengths 1 and extend branches
```{r}
Phylogeny_clustered_extended <- Phylogeny_clustered
NEdge <- length(Phylogeny_clustered_extended$edge.length)
edge_length <- replicate(NEdge, 1)
Phylogeny_clustered_extended$edge.length <- edge_length
N <- Ntip(Phylogeny_clustered_extended)
root_node <- N+1
root_to_tip <- dist.nodes(Phylogeny_clustered_extended)[1:N, root_node]
var(root_to_tip)
age_difference <- max(root_to_tip) - root_to_tip
tip_edges <- Phylogeny_clustered_extended$edge[, 2] <= Ntip(Phylogeny_clustered_extended)
Phylogeny_clustered_extended$edge.length[tip_edges] <- Phylogeny_clustered_extended$edge.length[tip_edges] + age_difference
is.ultrametric(Phylogeny_clustered_extended)
Phylogeny_clustered_extended <- rescaleRR(Phylogeny_clustered_extended, height = 1)

plot(Phylogeny_clustered_extended)

Phylogeny_random_extended <- Phylogeny_random
NEdge <- length(Phylogeny_random_extended$edge.length)
edge_length <- replicate(NEdge, 1)
Phylogeny_random_extended$edge.length <- edge_length
N <- Ntip(Phylogeny_random_extended)
root_node <- N+1
root_to_tip <- dist.nodes(Phylogeny_random_extended)[1:N, root_node]
var(root_to_tip)
age_difference <- max(root_to_tip) - root_to_tip
tip_edges <- Phylogeny_random_extended$edge[, 2] <= Ntip(Phylogeny_random_extended)
Phylogeny_random_extended$edge.length[tip_edges] <- Phylogeny_random_extended$edge.length[tip_edges] + age_difference
is.ultrametric(Phylogeny_random_extended)
Phylogeny_random_extended <- rescaleRR(Phylogeny_random_extended, height = 1)

plot(Phylogeny_random_extended)
```

#PhyloGLM 
```{r}
fit_random <- phyloglmstep(formula = formula_1, data = Growth_Form, phy = Phylogeny_random_extended, btol = 20)

fit_clustered <- phyloglmstep(formula = formula_1, data = Growth_Form, phy = Phylogeny_clustered_extended, btol = 20)
summary(fit_random)
summary(fit_clustered)
```
