---
title: "OMI with Envirem data"
author: "Marit Kuijt"
date: "2023-02-14"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/marit/OneDrive/Documenten/Masterstage Naturalis/brassicaceae-trait-climate-analysis")
```

```{r}
setwd("~/Masterstage Naturalis/brassicaceae-trait-climate-analysis")
library(geodata)
options(stringsAsFactors = FALSE)
library(raster)
library(tidyverse)
library(sp)
library(adehabitatMA)
library(optparse)
library(yaml)
library(dplyr)
library(ade4)
library(tibble)
library(logger)
library(CoordinateCleaner)
library(fuzzySim)
library(factoextra)
library(ggforce)
library(ggalt)
library(ggtree)
library(usdm)
library(subniche)
library(knitr)
library(phylolm)
library(ggnewscale)
library(ggplot2)
library(ggpubr)
library(ape)
library(phytools)
library(maptools)
library(Hmisc)
library(RRphylo)
```
#Read Data
##GIS data
```{r add_layers_ENVIREM}
# Read ENVIREM data and stack
remove(WorldENVIREM)
WorldENVIREM <- raster()

# The location of the TIFF file with the stacked layers described directly above
files.names <- list.files("./Data/ENVIREM")
# Turn the file names into layer names: strip the prefix (which might include
# the resolution) and strip the file extension
gis.layers.names <- files.names
gis.layers.names <- gsub("current_10arcmin_", "", gis.layers.names)

# Iterate over files
for (j in 1:length(files.names)) {
  
  # Stack with previously read layers
  WorldENVIREM <- stack(
    WorldENVIREM,
    
    # Read as raster
    raster(
      
      # Construct file name
      paste("./Data/ENVIREM/", files.names[j], sep = "")
    )
  )
}
```

##Extract coordinates from 1 layer
```{r}
WorldENVIREM1.spdf <- as(WorldENVIREM@layers[[1]], "SpatialPixelsDataFrame")
sp::proj4string(WorldENVIREM1.spdf) <- crs("+proj=longlat +ellps=WGS84")
coordinates <- WorldENVIREM1.spdf@coords
```

##UV Data
```{r}
UV_Data <- read.asciigrid("./Data/56459_UVB1_Annual_Mean_UV-B.asc")
plot(UV_Data)
UV_Data.raster <- raster(UV_Data)
sp::proj4string(UV_Data) <- crs("+proj=longlat +datum=WGS84")
```

###Extract coordinates from cells of ENVIREM and values from these cells of UV data
```{r}
UV_cells <- raster::cellFromXY(UV_Data.raster, coordinates)
UV_values <- raster::extract(UV_Data.raster, UV_cells)
UV_Bound <- cbind(coordinates, UV_values)
UV_Bound <- data.frame(UV_Bound)
sp::coordinates(UV_Bound) <- ~ x + y
sp::proj4string(UV_Bound) <- crs("+proj=longlat +ellps=WGS84")

UV_Bound.raster <- rasterFromXYZ(UV_Bound)
sp::proj4string(UV_Bound.raster) <- crs("+proj=longlat +ellps=WGS84")
names(UV_Bound.raster) <- "UV"
```

```{r}
soil_files <- list.files("./Data/Soil_Data_5m")
soil_files <- soil_files[-c(11, 12)]
WorldSoil <- raster()

for (i in 1:length(soil_files)) {
  Soil_Layer <- raster(paste("./Data/Soil_Data_5m/", soil_files[i], sep = ""))
  
  SoilCells <- raster::cellFromXY(Soil_Layer, coordinates)
  SoilValues <- raster::extract(Soil_Layer, SoilCells)
  SoilValues[is.na(SoilValues)] <- 0
  Soil_bound <- cbind(SoilValues, coordinates)
  Soil_bound <- data.frame(Soil_bound)

  sp::coordinates(Soil_bound) <- ~ x + y
  sp::proj4string(Soil_bound) <- crs("+proj=longlat +ellps=WGS84")

  Soil_bound.raster <- rasterFromXYZ(Soil_bound)
  sp::proj4string(Soil_bound.raster) <- crs("+proj=longlat +ellps=WGS84")
  names(Soil_bound.raster) <- soil_files[i]
  WorldSoil <- stack(WorldSoil, Soil_bound.raster)
}
plot(WorldSoil)
```



```{r}
WorldENVIREM <- crop(WorldENVIREM, UV_Bound.raster)
WorldENVIREM <- stack(WorldENVIREM, UV_Bound.raster)
WorldENVIREM <- stack(WorldENVIREM, WorldSoil)
plot(WorldENVIREM)

rm(gis.layers.names, files.names, j)
```


##BioClim Data
```{r add_layers_World_clim}
## Read Bioclim data and stack
remove(WorldBioClim)
WorldBioClim <- raster()

# The location of the TIFF file with the stacked layers described directly above
files.names <- list.files("./Data/wc2.1_10m")
# Turn the file names into layer names: strip the prefix (which might include
# the resolution) and strip the file extension
gis.layers.names <- files.names
gis.layers.names <- gsub('./Data/wc2.1_10m_','',gis.layers.names)
gis.layers.names <- gsub('.tif','',gis.layers.names)

gis.layers <- c()
# Iterate over files
for (i in 1:length(files.names)) {
  
  # Stack with previously read layers
  BioClimLayer <- raster(paste("./Data/wc2.1_10m/", files.names[i], sep = ""))
  BioClimLayer <- crop(BioClimLayer, WorldENVIREM)
  WorldBioClim <- stack(
    WorldBioClim, BioClimLayer
  )
}

rm(gis.layers.names, files.names, i)
```

##Stack all BioClimatic Layers together
```{r}
All_Layers <- stack(WorldBioClim, WorldENVIREM)
All_Layers.spdf <- as(All_Layers, "SpatialPixelsDataFrame")
sp::proj4string(All_Layers.spdf) <- crs("+proj=longlat +ellps=WGS84")
```


##Occurrence data
```{r}
occurrences.spdf <- new(
    "SpatialPointsDataFrame", 
    coords = structure(
        numeric(0), 
        .Dim = c(0L, 2L),
        .Dimnames = list( NULL, c("x", "y") )
    ),  
    bbox = structure(
        c(1,1,1,1), 
        .Dim = c(2L, 2L),                         
        .Dimnames = list( c("x","y"), c("min","max") )
    )
) 
sp::proj4string(occurrences.spdf) <- crs("+proj=longlat +ellps=WGS84")

Tribe <- "Alysseae"

# Populate the empty dataframe with lat/lon values from the taxa list
gbif_summary <- read.csv("Data/gbif_summary.csv")
# set for example i in 1:100 in the case of memory limits 
taxa.names <- gbif_summary[gbif_summary$Tribe == Tribe,]
taxa.names <- taxa.names$Name[taxa.names$BrassiBase_status == "accepted" & taxa.names$gbif_records > 20]
for (j in 1:length(taxa.names)) {
    
    # Prevent out of bounds errors
    if (j > length(taxa.names)) {
        break
    }
  
	# Read occurrences
    species.key <- gbif_summary$speciesKey[gbif_summary$Name == taxa.names[j] & !duplicated(gbif_summary$Name)]
    csv.file <- paste("Data/gbif_occurrences_per_key/", species.key, ".csv", sep = "")
    csv.data <- read.csv(csv.file)
    #tryCatch(csv.data <- read.csv(csv.file), error=function(e) force(do.next))
    points.df <- dplyr::select(
        csv.data,
        decimalLongitude,
        decimalLatitude,
        species
    )
    #Coordinate Cleaner
    points.df <- na.omit(points.df)
    if(nrow(points.df) < 20 & file.exists(paste("Data/gbif_occurrences_per_key/", species.key, "_extra.csv", sep = "")) == T){
    csv.file.extra <- paste("Data/gbif_occurrences_per_key/", species.key, "_extra.csv", sep = "")
    csv.data.extra <- read.csv(csv.file.extra)
    points.extra.df <- dplyr::select(
    csv.data.extra,
    decimalLongitude,
    decimalLatitude,
    species
    )    
    points.extra.df <- na.omit(points.extra.df)
    points.df <- rbind(points.df, points.extra.df)}
    
        if (nrow(points.df) < 20) {next}
    else {
    
    #points.df <- points.df[rowSums(points.df[,1:2]) > 0,]
    
    points.df <- cc_cap(points.df, lon = "decimalLongitude", lat = "decimalLatitude", species = "species")
    points.df <- cc_cen(points.df, lon = "decimalLongitude", lat = "decimalLatitude", species = "species")
    points.df <- cc_dupl(points.df, lon = "decimalLongitude", lat = "decimalLatitude", species = "species")
    points.df <- cc_equ(points.df, lon = "decimalLongitude", lat = "decimalLatitude")
    points.df <- cc_gbif(points.df, lon = "decimalLongitude", lat = "decimalLatitude", species = "species")
    points.df <- cc_inst(points.df, lon = "decimalLongitude", lat = "decimalLatitude", species = "species")
    points.df <- cc_outl(points.df, lon = "decimalLongitude", lat = "decimalLatitude", species = "species")
    points.df <- cc_sea(points.df, lon = "decimalLongitude", lat = "decimalLatitude")
    
    if (nrow(points.df) == 0) {next}
    points.df <- cc_zero(points.df, lon = "decimalLongitude", lat = "decimalLatitude")
    
    if (nrow(points.df) < 20) {next}
    sp::coordinates(points.df) <- ~ decimalLongitude + decimalLatitude
    sp::proj4string(points.df) <- crs("+proj=longlat +ellps=WGS84")
    
    # Populate SpatialPointsDataFrame for focal taxon
    focal.spdf <- SpatialPointsDataFrame(
    	points.df, 
    	data.frame(species = rep(taxa.names[j], NROW(points.df)))
    )
    sp::proj4string(focal.spdf) <- crs("+proj=longlat +ellps=WGS84")
    
    # Append to cumulative data frame
    occurrences.spdf <- rbind(occurrences.spdf, focal.spdf)
    
    }
}
```


###Plot occurrences
```{r}
occurrences.df <- data.frame(occurrences.spdf@data)
occurrences.df <- cbind(occurrences.df, occurrences.spdf@coords)
lon <- dplyr::select(occurrences.df, x)
lat <- dplyr::select(occurrences.df, y)

data("wrld_simpl")
shape <- wrld_simpl
shape_df <- fortify(shape, region = 'ISO3')

ggplot() +
  geom_polygon(data = shape_df, aes(x = long, y = lat, group = group), fill = "white", color = "black")+
  geom_point(aes(x, y, fill = species), data = occurrences.df, shape = 21, size = 2, alpha = 0.5, col = "black") +
    coord_equal(xlim = c((min(lon) - 1), (max(lon) + 1)), ylim = c(min(lat) - 1, max(lat) + 1)) +
  xlab("Longitude") + ylab("Latitude") +
  theme(panel.border = element_rect(fill=NA), legend.position = 'none')
```


```{r}
names <- unique(occurrences.spdf$species)
```

##GIS data per cell (na.omit to exclude areas without data in oceans etc.)
```{r}
traitvalues.df <- slot(All_Layers.spdf, "data")
traitvalues.df <- tibble::rowid_to_column(traitvalues.df, "ID")
traitvalues.df <- na.omit(traitvalues.df)
```

##Presence/absence matrix per cell
```{r}
npoints <- adehabitatMA::count.points(occurrences.spdf, All_Layers.spdf)
noccurrences.df <- slot(npoints, "data")
noccurrences.df <- tibble::rowid_to_column(noccurrences.df, "ID")
noccurrences.df <- subset(noccurrences.df, noccurrences.df$ID %in% traitvalues.df$ID)
noccurrences.df <- subset(noccurrences.df, select = -ID)
noccurrences.df[noccurrences.df > 0] <- 1
noccurrences.df <- noccurrences.df[rowSums(noccurrences.df[]) > 0,]
```

##Select only cells with presence for noccurrences.df and traitvalues.df
```{r}
noccurrences.df$ID <- rownames(noccurrences.df)
traitvalues.df <- subset(traitvalues.df, traitvalues.df$ID %in% noccurrences.df$ID)
noccurrences.df <- subset(noccurrences.df, select = -ID)
```


#OMI analysis
##First a PCA
```{r}
traitvalues.dudi <- ade4::dudi.pca(
    traitvalues.df[2:length(traitvalues.df)], 
    scannf = F
)
```

##Then the OMI analysis
```{r}
niche.dudi <- ade4::niche(traitvalues.dudi, noccurrences.df, scannf = F)
par(mfrow = c(1,1))
scatter(niche.dudi)
screeplot(niche.dudi)
kable(niche.param(niche.dudi))
ordered1 <- order(niche.dudi$c1$CS1)
var1 <- rownames(niche.dudi$c1[ordered1[1],])
ordered2 <- order(niche.dudi$c1$CS2)
var2 <- rownames(niche.dudi$c1[ordered2[1],])
```

```{r}
rbind(
  Eigenvalue = niche.dudi$eig,
  Proportion = niche.dudi$eig/sum(niche.dudi$eig),
  Cumulative = cumsum(niche.dudi$eig)/sum(niche.dudi$eig))
```

##Save output
```{r}
write.csv(niche.dudi$tab, file = "./Results/niche_traits_merged_Lepidium", append = T, quote = F )
```


#Read Tree and filter to species in OMI
```{r}
MetaData1 <- read_csv("./Data/2e.metadata.csv")
BrassiToL <- read.tree("Data/species_tree.tre")
names_to_drop <- setdiff(MetaData1$Name, names)
tips <- MetaData1$Library_ID[MetaData1$Name %in% names_to_drop]
BrassiToL_Lepidium <- drop.tip(BrassiToL, tips)
```

#Combine trait data of species and enviroment
```{r}
Growth_Form_ID <- subset(MetaData1, Tribe = Tribe, select = c(Library_ID, Growth_habit, Name))
Growth_Form <- data.frame(Growth_Form_ID)
rownames(Growth_Form) <- Growth_Form_ID$Library_ID
Growth_Form <- subset(Growth_Form, select = c(Growth_habit, Name))
```

```{r}
Growth_Form$Growth_habit[Growth_Form$Growth_habit == "H"] <- 0
Growth_Form$Growth_habit[Growth_Form$Growth_habit == "W"] <- 1
Growth_Form$Growth_habit <- base::as.numeric(Growth_Form$Growth_habit)
```

```{r}
MeanBioClim <- niche.dudi$tab
rownames(MeanBioClim) <- gsub(".", " ", rownames(MeanBioClim), fixed = T)
common1 <- intersect(rownames(MeanBioClim), Growth_Form_ID$Name)
MeanBioClim <- MeanBioClim[common1,]
SpeciesNames <- rownames(MeanBioClim)
```

```{r}
common2 <- intersect(rownames(Growth_Form), BrassiToL_Lepidium$tip.label)
Growth_Form <- Growth_Form[common2,]
Growth_Form <- distinct(Growth_Form, Name, .keep_all = T)
```

##Add the ENVIREM OMI-transformed climatic data for each species
```{r}
Growth_Form[,3:48] <- NA
Trait_names <- colnames(MeanBioClim)
colnames(Growth_Form)[3:48] <- Trait_names
Growth_Form <- Growth_Form[,-49]
for (i in 1:length(Trait_names)) {
  for(j in 1:length(SpeciesNames)){
    Growth_Form[Growth_Form$Name == SpeciesNames[j], colnames(Growth_Form) == Trait_names[i]] <- MeanBioClim[rownames(MeanBioClim) == SpeciesNames[j], colnames(MeanBioClim) == Trait_names[i]]
  }
}
```

```{r}
Growth_Form <- na.omit(Growth_Form)
Growth_Form <- Growth_Form[,c(1:3, 14:21, 4:13, 22:48)]
```

```{r}
Growth_Form_Name <- subset(Growth_Form, select = Name)
Growth_Form_Trait <- subset(Growth_Form, select = Growth_habit)
  Growth_Form_Trait$Growth_habit <- as.factor(Growth_Form_Trait$Growth_habit)
Growth_Form_annotation <- subset(Growth_Form, select = -c(1, 2))
```

```{r}
Growth_Form$Growth_habit <- factor(Growth_Form$Growth_habit, levels = 0:1, labels = c(0,1))
```


#Boxplots
##Make boxplots per variable ENVIREM

```{r}
Growth_Form_Long <- Growth_Form
Growth_Form_Long$ID <- rownames(Growth_Form_Long)
Growth_Form_Long <- Growth_Form_Long[,c(1:2, 39, 3:49)]
Growth_Form_Long <- gather(Growth_Form_Long, BCVariable, Value, 4:49, factor_key = T)

boxplot <- ggplot(Growth_Form_Long, aes(Growth_habit, Value, group = Growth_habit, fill = Growth_habit)) +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y.npc = "bottom") +
facet_wrap(~ BCVariable)
boxplot
```

#Make a dendrogram of all variables to see the relationship between them

```{r}
MeanBioClim.matrix <- as.matrix(MeanBioClim)
v <- varclus(MeanBioClim.matrix)
h <- v$hclust
hd <- as.dendrogram(h)
dend <- fviz_dend(hd, k = 5, cex=0.8, palette = "jco", color_labels_by_k = T, 
      rect = T, rect_lty = 0, rect_border = "jco", rect_fill = T)
dend
vars <- dend$layers[[2]]$data
clusters <- dend$layers[[4]]$data

colours <- unique(vars$col)
randomvars <- data.frame()

for (i in 1:length(colours)) {
showvar <- vars[sample(which(vars$col == colours[i]), 1), ]
randomvars <- rbind(randomvars, showvar)
}
randomvars$label
```

#Make brench lengths 1 and extend branches
```{r}
BrassiToL_Lepidium_extended <- BrassiToL_Lepidium
NEdge <- length(BrassiToL_Lepidium_extended$edge.length)
edge_length <- replicate(NEdge, 1)
BrassiToL_Lepidium_extended$edge.length <- edge_length
N <- Ntip(BrassiToL_Lepidium_extended)
root_node <- N+1
root_to_tip <- dist.nodes(BrassiToL_Lepidium_extended)[1:N, root_node]
var(root_to_tip)
age_difference <- max(root_to_tip) - root_to_tip
tip_edges <- BrassiToL_Lepidium_extended$edge[, 2] <= Ntip(BrassiToL_Lepidium_extended)
BrassiToL_Lepidium_extended$edge.length[tip_edges] <- BrassiToL_Lepidium_extended$edge.length[tip_edges] + age_difference
is.ultrametric(BrassiToL_Lepidium_extended)
BrassiToL_Lepidium_extended <- rescaleRR(BrassiToL_Lepidium_extended, height = 1)

```

#Plot tree
```{r}
p <- ggtree(BrassiToL_Lepidium_extended) %<+% MetaData1 +
  geom_tiplab(size = 3, offset = 0.2) +
  geom_tiplab(aes(label = Name), offset = 0.5, size = 3) + 
  geom_text(aes(label=node), hjust=-.3)
p2 <- gheatmap(p, Growth_Form_Trait, offset = 1.5, width = 0.5)
p3 <- p2 + new_scale_fill()
gheatmap(p3, Growth_Form_annotation, offset = 2, width = 5, colnames_angle = -45, colnames_offset_y = 5, colnames_position = "top") + scale_fill_continuous(na.value = "white")
```

#Pagel's Lambda 
```{r}
Pagel <- data.frame(row.names = Trait_names)
Pagel$lambda <- NA
for (i in 1:length(Trait_names)) {
trait <- Growth_Form[,i+2]
names(trait) <- rownames(Growth_Form)
lambda <- phylosig(BrassiToL_Lepidium_extended, trait, method="lambda", test=TRUE, nsim=999)
Pagel$lambda[i] <- lambda$lambda
}
```

#Correlation test
```{r}
correlation <- cor(Growth_Form_annotation, method="spearman")
correlation <- as.data.frame(correlation)
col_1 <- which(correlation$current_10arcmin_PETColdestQuarter < 0.7 & correlation$current_10arcmin_PETColdestQuarter > -0.7)
col_2 <- which(correlation$current_10arcmin_embergerQ < 0.7 & correlation$current_10arcmin_embergerQ > -0.7)
Trait_no_col <- intersect(col_1, col_2)
Trait_names <- rownames(correlation)
Trait_names[Trait_no_col]
```

```{r}
vif <- vif(MeanBioClim[c(2, 3, 5, 30, 31, 45)])
vif
```

#Make a model to use in phyloglm based on the boxplots and vif
```{r}
unique(Growth_Form$Name)

formula_0 <- Growth_habit ~  wc2.1_10m_bio_10 + wc2.1_10m_bio_11 + wc2.1_10m_bio_13 + current_10arcmin_PETColdestQuarter + current_10arcmin_PETDriestQuarter + ROOTS5min.nc

formula_1 <- Growth_habit ~ wc2.1_10m_bio_1 + wc2.1_10m_bio_9 + wc2.1_10m_bio_10 + wc2.1_10m_bio_12 + wc2.1_10m_bio_13 + wc2.1_10m_bio_14 + wc2.1_10m_bio_16 + wc2.1_10m_bio_17 + wc2.1_10m_bio_18 + wc2.1_10m_bio_19 + current_10arcmin_annualPET + current_10arcmin_aridityIndexThornthwaite + current_10arcmin_climaticMoistureIndex + current_10arcmin_embergerQ + current_10arcmin_growingDegDays0 + current_10arcmin_growingDegDays5 + current_10arcmin_minTempWarmest + current_10arcmin_monthCountByTemp10 + current_10arcmin_PETDriestQuarter + current_10arcmin_PETWarmestQuarter + current_10arcmin_thermicityIndex

formula_2 <- Growth_habit ~ wc2.1_10m_bio_13 + wc2.1_10m_bio_17 + wc2.1_10m_bio_19 + current_10arcmin_minTempWarmest + current_10arcmin_monthCountByTemp10 + current_10arcmin_PETDriestQuarter + current_10arcmin_PETseasonality

formula_3 <- Growth_habit ~ current_10arcmin_climaticMoistureIndex


```

#PhyloGLM 
```{r}
fit <- phyloglmstep(formula = formula_0, data = Growth_Form, phy = BrassiToL_Lepidium_extended, method="logistic_MPLE", direction = "both", btol = 50, log.alpha.bound = 8, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = TRUE, k=2)
summary(fit)
```